Double Standards in Judging Collective Action

### 2022-08-22 Documents/manuscript.Rmd
update manuscript

---
title: "Double Standards in Judging Collective Action"
header: "DOUBLE STANDARDS"
keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change. In four studies (total *N* = 2,246), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*). In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action. In two preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police in the United States. We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members. Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: preprint-template.tex

@@

In this paper, we examine double standards in judging collective action---that is, whether observers judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (*identity-based double standards*) or on the protesters' cause and how it aligns with their own ideological positions (*ideology-based double standards*). In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action. In two preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police in the United States. We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own (system-challenging) ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.

@@

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1L,
        Participant != Protesters ~ -1L,
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1L,
        Participant == "White" & Cause == "Against" ~ +1L,
        TRUE ~ -1L,
      ),
      h2a_contrast = case_when(
        Cause == "For" ~ -1L,
        Cause == "Against" ~ +1L
      ),
      nh1_contrast = case_when(
        Participant == "Black" ~ -1L,
        Participant == "White" ~ +1L,
      ),
      nh2_contrast = case_when(
        Participant == "White" & Cause == "Against" ~ -1L,
        Participant == "White" & Cause == "For" ~ +1L,
        Participant == "Black" ~ 0L
      ),
      nh3_contrast = case_when(
        Participant == "Black" & Cause == "Against" ~ -1L,
        Participant == "Black" & Cause == "For" ~ +1L,
        Participant == "White" ~ 0L
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    group_by(.draw) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~mean(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

@@

  # Define contrasts
  m5 <- m5 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1L,
        Participant != Protesters ~ -1L,
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1L,
        Participant == "White" & Cause == "Against" ~ +1L,
        TRUE ~ -1L,
      )
    )

  # Calculate contrasts
  m5 <- m5 %>% 
    group_by(.draw) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~mean(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

@@

We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses. To that end, we extended Model 1 by estimating participants' responses as a function of their *z*-standardized identification with their racial ingroup. As in Model 2, we modeled this relationship with a fixed and a varying effect. We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast"])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable  than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast"])`). Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgements about collective action by ingroup and outgroup members.

@@

@@

Our research was motivated by real-world examples that seemed to suggest that where people draw the line between acceptable and unacceptable protest actions depends on *who* the protesters are and *what* they are protesting---in other words, that there are double standards in judging collective action. Our findings, however, showed that it is not so simple. Contradicting our hypothesis of identity-based double standards, participants in two preregistered experiments did not show ingroup bias when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests. Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position. 
<!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.  -->
In the remainder of this General Discussion, we discuss theoretical and practical implications for understanding the often divided response to collective action and consider some limiting conditions on our findings.

@@

Our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause. Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values. In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder. By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provided a paradigm that we hope will stimulate research on when and why people differ in making this distinction.

After all, this distinction can be consequential. If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism. In this way, our findings could explain why 45% of Republicans (2% of Democrats) supported the storming of the U.S. Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent. If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent. In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022]. For those reasons, researchers should further investigate when and why people apply different standards when judging collective action for various causes.

Despite its strengths, our research is qualified by several limitations. First, our research examined the hypothesized relationships for only two political causes---protecting workers' rights, defunding the police---and thus provided only limited evidence that our findings generalize beyond those causes. Relying on few stimuli to establish an effect threatens the validity and replicability of research findings and severely constrains psychological research generalize beyond those few stimuli [@yarkoni_generalizability_2022]. Future research should sample a wider range of causes to address this pervasive but often ignored problem [@judd_treating_2012]. This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes. Second, our research was based on two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) samples, which severely limits the cross-cultural generalizability of our findings. Future research should include samples from a wider range of cultural context to examine how the observed relationships vary across cultural contexts.

@@

>>>>

---
title: "Double Standards in Judging Collective Action"
header: "DOUBLE STANDARDS"
keywords: protest, collective action, double standards, social identity theory, system justification theory, Black Lives Matter
abstract: "Collective action is a powerful force driving social change but often sparks contention about what actions are acceptable means to effect social change. In five studies (total *N* = 2,979), we investigated double standards in judging collective action---that is, whether observers will judge the same protest actions as more acceptable if the protesters are ingroup members and their cause aligns with the ingroup's interests (*identity-based double standards*) or if the protesters' cause aligns with the observers' ideological positions (*ideology-based double standards*). In two studies, we developed an instrument of 25 controversial protest actions, based in item response theory, to measure where people draw the line between acceptable and unacceptable forms of collective action. In three preregistered experiments, we used this instrument to test for double standards in judging collective action for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States. We found evidence for ideology-based, but not for identity-based, double standards: Participants judged the same protest actions to be more acceptable when the protesters' cause aligned with their own ideological position---but showed no ingroup bias when judging collective action by ingroup and outgroup members. Our findings have theoretical and practical implications for understanding the often divided response to prominent social movements."
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: preprint-template.tex

@@

In this paper, we examine double standards in judging collective action---that is, whether observers judge the same protest actions as more or less acceptable depending on their own and the protesters' group memberships (*identity-based double standards*) or on the protesters' cause and how it aligns with their own ideological positions (*ideology-based double standards*). In two studies, we first develop an instrument of 25 controversial protest actions, based in item response theory, to capture double standards in judging collective action. In three preregistered experiments, we then use this instrument to test for double standards in judging protest actions for workers' rights in the United Kingdom and for or against defunding the police and for or against restricting abortion in the United States. We find evidence for ideology-based, but not for identity-based, double standards: People judge the same protest actions to be more acceptable when the protesters' cause aligns with their own ideological positions---but do not show ingroup bias when judging collective action by ingroup and outgroup members.

@@

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1/8,
        Participant != Protesters ~ -1/8,
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1/8,
        Participant == "White" & Cause == "Against" ~ +1/8,
        TRUE ~ -1/8,
      ),
      h2a_contrast = case_when(
        Cause == "For" ~ -1/8,
        Cause == "Against" ~ +1/8
      ),
      nh1_contrast = case_when(
        Participant == "Black" ~ -1/8,
        Participant == "White" ~ +1/8
      ),
      nh2_contrast = case_when(
        Participant == "White" & Cause == "Against" ~ -1/4,
        Participant == "White" & Cause == "For" ~ +1/4,
        Participant == "Black" ~ 0
      ),
      nh3_contrast = case_when(
        Participant == "Black" & Cause == "Against" ~ -1/4,
        Participant == "Black" & Cause == "For" ~ +1/4,
        Participant == "White" ~ 0
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    group_by(.draw) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~sum(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

@@

  # Define contrasts
  m5 <- m5 %>% 
    mutate(
      h1a_contrast = case_when(
        Participant == Protesters ~ +1/8,
        Participant != Protesters ~ -1/8
      ),
      h1b_contrast = case_when(
        Participant == "Black" & Cause == "For" ~ +1/8,
        Participant == "White" & Cause == "Against" ~ +1/8,
        TRUE ~ -1/8,
      )
    )

  # Calculate contrasts
  m5 <- m5 %>% 
    group_by(.draw, x_ide) %>% 
    summarize(
      across(
        ends_with("_contrast"),
        ~mean(z * .)
      )
    ) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(x_ide, contrast) %>% 
    median_qi(d)

@@

We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses. To that end, we extended Model 1 by estimating participants' responses as a function of their *z*-standardized identification with their racial ingroup. As in Model 2, we modeled this relationship with a fixed and a varying effect. We found, however, that even participants who strongly identified with their racial ingroup (+1*SD*) did not consider protest actions performed by their ingroup to be more acceptable than the same actions performed by the relevant outgroup (`r with(m5, d_text[contrast == "h1a_contrast" & x_ide == 1L])`) and did not consider protest actions for a cause that was aligned with their ingroup's interests to be more acceptable than protest actions for a cause not aligned with their ingroup's interests (`r with(m5, d_text[contrast == "h1b_contrast" & x_ide == 1L])`). Our non-preregistered analyses thus suggested that group identification did not moderate group differences in judgements about collective action by ingroup and outgroup members.

@@

# Experiment 3

We preregistered the sample size, hypotheses, inclusion/exclusion criteria, measures, and manipulations ([https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822](https://osf.io/kphyw/?view_only=59f5792a6d3f4fa48516ea9d5f637822)). We deviated from the preregistration in two ways. First, we stopped collecting data before reaching the preregistered sample size. We did so as the Supreme Court decision on reproductive rights, announced on June 24, 2022, had rendered the experimental manipulation invalid. Second, we do not report results for two further hypotheses---testing moral conviction as a potential mechanism underlying ideology-based double standards---as those findings are beyond the scope of this manuscript and will be reported in a future publication. We made all materials, data, and analysis scripts available online ([https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62](https://osf.io/d3yev/?view_only=40782034017c40f0bcecb1cc87760b62)).

We used a 2 (quasi-experimental: conservative/liberal participants) $\times$ 2 (quasi-experimental: male/female participants) $\times$ 2 (experimental: for/against restricting abortion) between-subjects design to test our hypotheses.

We recruited 804 participants from the Prolific subject pool who were 18 years old or older and lived in the United States.^[Data were collected between June 10 and 23, 2022. Participants received, on average, \$24.70 per hour of participation.] We balanced the sample's gender and political composition so that half of all participants identified as conservative and the other half identified as liberal and so that roughly half of the participants in each group were women and men. As preregistered, we excluded 71 participants who failed at least one of five attention checks or reported a different political orientation than they had reported in the Prolific prescreening questionnaire. As noted above, we had to stop collecting data before reaching the preregistered sample size of 800 eligible participants. This resulted in a final sample of 733 participants ($\textit{Mdn} = 37$ years, age range: 18--93 years) of whom 181 were conservative men, 190 were conservative women, 185 were liberal men, and 177 were liberal women.

After answering where they would you place themselves along the political spectrum, in line with their response in the Prolific prescreening questionnaire, participants read the following paragraph:

> Recent news suggested that the Supreme Court might soon overturn its *Roe v. Wade* decision that, so far, has guaranteed legal abortions in the first 20--24 weeks of pregnancy. Overturning *Roe v. Wade* would allow states to enact laws to further restrict or ban abortions.

\noindent In the supporting abortion restrictions condition, participants then read a text presenting arguments for restricting or banning abortion:

> Some people welcome this news because they oppose legal abortions.  They argue that life begins at conception and that, therefore, abortion ends the life of an unborn child.  As human life is sacred, ending an unborn life is wrong and cannot be justified by a pregnant person's right to choose.  In this view, restricting or banning abortions is the only way to protect unborn lives. We refer to this position as "supporting abortion restrictions".

\noindent In the opposing abortion restrictions condition, participants instead read a text presenting arguments against restricting or banning abortion:

> Some people are alarmed by this news because they support legal abortions. They argue that pregnant people have a right to control their own body and, therefore, have a right to safe and legal abortions. Restricting or banning abortion would lead to more illegal and unsafe abortions. In this view, providing legal abortions is the only way to protect pregnant people's rights and to prevent harm. We refer to this position as "opposing abortion restrictions".

\noindent Participants were then asked to select which position the text described and what arguments for that position were presented in the text. If they selected the wrong answer, they were instructed to reread the text and select the right answer. Participants then rated to what extent they support or oppose restricting abortion and to what extent their feelings about this issue are a moral conviction.

Participants then read about protesters protesting either for (in the supporting abortion restrictions condition) or against (in the opposing abortion restrictions condition) restricting or banning abortion:

> Earlier, you read why some people [support/oppose] further restricting or banning abortion. Some local residents who hold this view want to protest [for/against] restricting or banning abortion.

\noindent Participants rated which of several protest actions they consider acceptable means to protest for or against restricting abortion. On the next pages, participants completed the remaining measures. On the final page, participants responded to two attention checks that assessed whether they paid attention to the experimental manipulation: "In this study, you read about some people's view on a policy change. What was the policy change?" and "In this study, you then read about some protesters. Were the protesters for or against restricting abortion?".

We measured the outcome variable by asking participants to think about the protesters they had read about and to decide, for each of 25 protest actions presented in a randomized order, whether they thought this action was "an acceptable means for them to protest [for/against] restricting or banning abortion" (1 = *yes*, 0 = *no*; see Table C1).

We included additional measures to describe the sample, to report as descriptive statistics, to use in non-preregistered analyses or future analyses not reported in this paper.

We measured support for restricting abortion with one item: "Do you support or oppose restricting (or banning) abortion?" (1 = *strongly oppose*, 5 = *strongly support*). We also included a  three-item measure of moral conviction [@ryan_reconsidering_2014] that assessed to what extent participants' feelings about restricting or banning abortion were, for example, connected to their core moral beliefs or convictions (1 = *not at all*, 5 = *very much*).

We measured gender-related system-justifying beliefs with eight items [adapted from @jost_exposure_2005], for example, "in general, relations between men and women are fair" (1 = *strongly disagree*, 7 = *strongly agree*; McDonald's $\omega = .93$). A confirmatory factor analysis model in which all items loaded onto a single factor showed acceptable fit, $\chi^2 (20) = 111.90; \text{CFI} = 0.97; \text{TLI} = 0.95; \text{RMSEA} = 0.09, [0.08, 0.11]$.

We will measured moral concerns with the 36-item moral foundations questionnaire [MFQ-2, @atari_morality_2022] which assesses to what extent participants endorse concerns about care, equality, proportionality, loyalty, authority, and purity (e.g., "I believe chastity is an important virtue."; 1 = *does not describe me at all*, 5 = *describes me extremely well*). We embedded three further attention checks within the questionnaire (e.g., "To show that you are paying attention and giving your best effort, please select 'moderately describes me'.").

In addition to demographic questions, we asked participants to what extent they identified with their gender (1 = *strongly disagree*, 7 = *strongly agree*) and measured political orientation with a one-item semantic differential scale: "People often describe their political orientation as liberal or conservative. On a scale from liberal to conservative, where would you position yourself?" (1 = *liberal*, 7 = *conservative*).

As expected, participants who identified as conservative reported being far more conservative ($M = 6.05, SD = 0.79$) than participants who identified as liberal ($M = 1.73, SD = 0.80; \text{Cohen's}~d = 5.43$). Likewise, conservative participants tended to endorse system-justifying beliefs ($M = 5.10, SD = 0.95$) while liberal participants tended to reject system-justifying beliefs ($M = 2.86, SD = 1.03; d = 2.26$). As expected, conservative and liberal participants differed in their reactions to the manipulation as the former tended to support restricting abortion ($M = 3.77, SD = 1.48$) while the latter tended to oppose restricting abortion ($M = 1.28, SD = 0.85; d = 2.06$). When comparing politically balanced subsamples of men and women, women ($M = 2.55, SD = 1.76$) did not differ from men ($M = 2.55, SD = 1.72$) in their attitudes toward restricting abortion ($d = -0.02$).

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Results from the preregistered analyses for Experiment 3}
\includegraphics[scale=1]{../Experiment 3/figures/figure-6}
\caption*{\textit{Note.} Results show the estimated \textit{z}-standardized tendency ($\theta_{(z)}$) to consider more controversial actions acceptable means to protest for or against restricting abortion as a function of the participants' ideology and gender.}
\label{fig:f6}
\end{figure*}
```

  # THIS MIGHT BE WRONG

  # Import results
  m1 <- read_rds("../Experiment 3/results/m1_post.rds")

  # Define contrasts
  m1 <- m1 %>% 
    mutate(
      h1_contrast = case_when(
        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      h2a_contrast = case_when(
        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      h2b_contrast = case_when(
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L
      ),
      nh1_contrast = case_when(
        participant_ideology == "Conservative" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh2_contrast = case_when(
        participant_ideology == "Liberal" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      ),
      nh3_contrast = case_when(
        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & participant_gender == "Man" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh4_contrast = case_when(
        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "For" ~ +1L,
        participant_ideology == "Conservative" & participant_gender == "Woman" & protesters_cause == "Against" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh5_contrast = case_when(
        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      ),
      nh6_contrast = case_when(
        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
        participant_ideology == "Liberal" & participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        TRUE ~ NA_integer_
      )
    )

  # Calculate contrasts
  m1 <- m1 %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "direction",
      values_drop_na = TRUE
    ) %>% 
    group_by(.draw, contrast, direction) %>% 
    summarize(z = mean(z)) %>% 
    summarize(d = sum(z * direction)) %>% 
    pivot_wider(
      names_from = contrast,
      values_from = d
    ) %>% 
    mutate(nh7_contrast = nh2_contrast - nh1_contrast) %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "d"
    ) %>% 
    group_by(contrast) %>% 
    median_qi(d)

To test our hypotheses, we used a two-parameter logistic item response theory model, identical to the one used in Experiment 2, to estimate how likely participants were to consider each action an acceptable means of protest as a function of the protesters' cause and of the participants' gender and ideology. Figure \ref{fig:f6} shows the results of our preregistered analyses.

As in Experiments 1 and 2, we found no evidence for identity-based double standards: Female participants did not consider protest actions *against* restricting abortion to be any more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(m1, d_text[contrast == "h1_contrast"])`). Instead, we again found evidence for ideology-based double standards: Conservative participants considered the same protest actions to be more acceptable, on average, when protesters *supported* restricting abortion (`r with(m1, d_text[contrast == "nh1_contrast"])`) while liberal participants considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion (`r with(m1, d_text[contrast == "nh2_contrast"])`).
This pattern was consistent across subsamples as conservative women (`r with(m1, d_text[contrast == "nh4_contrast"])`) and, to a lesser extent, conservative men (`r with(m1, d_text[contrast == "nh3_contrast"])`) considered the same protest actions to be more acceptable when protesters *supported* restricting abortion and as both liberal men (`r with(m1, d_text[contrast == "nh5_contrast"])`) and women (`r with(m1, d_text[contrast == "nh6_contrast"])`) considered the same protest actions to be more acceptable when protesters *opposed* restricting abortion. This double standard was, however, more pronounced among liberal participants (`r with(m1, d_text[contrast == "nh7_contrast"])`). In this way, Experiment 3 provided evidence for both ideological symmetry and asymmetry in judging collective action for and against restricting abortion.

  # Import results
  ide <- read_rds("../Experiment 3/results/ide_post.rds")

  # Define contrasts
  ide <- ide %>% 
    mutate(
      h1_contrast = case_when(
        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        TRUE ~ -1L,
      ),
      nh1_contrast = case_when(
        participant_gender == "Woman" & protesters_cause == "Against" ~ +1L,
        participant_gender == "Woman" & protesters_cause == "For" ~ -1L,
        TRUE ~ NA_integer_
      ),
      nh2_contrast = case_when(
        participant_gender == "Man" & protesters_cause == "Against" ~ +1L,
        participant_gender == "Man" & protesters_cause == "For" ~ -1L,
        TRUE ~ NA_integer_
      )
    )

  # Calculate contrasts
  ide <- ide %>% 
    pivot_longer(
      ends_with("_contrast"),
      names_to = "contrast",
      values_to = "direction",
      values_drop_na = TRUE
    ) %>% 
    group_by(.draw, z_ide, contrast, direction) %>% 
    summarize(z = mean(z)) %>% 
    summarize(d = sum(z * direction)) %>% 
    group_by(z_ide, contrast) %>% 
    median_qi(d)

  # Make text
  ide <- ide %>% 
    mutate(
      across(
        c(d, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      d_text = glue(
        "$d = {d}, [{.lower}, {.upper}]$"
      )
    )

<!-- We explored whether group identification moderated how the participants' and the protesters' group memberships affected participants' responses. We found some evidence that, when considering only strongly identified (+1*SD*) men and women, female participants considered protest actions *against* restricting abortion to be more acceptable, on average, than male participants or protest actions *for* restricting abortion (`r with(ide, d_text[contrast == "h1_contrast" & z_ide == 1L])`). Our non-preregistered analyses thus suggested that gender identification might moderate gender differences in judgements about collective action concerning reproductive rights. -->

  # Import results
  sjb <- read_rds("../Experiment 3/results/sjb_post.rds")

  # Hypothesis 2
  sjb <- sjb %>% 
    distinct(.draw, b_sjb_for, b_sjb_against) %>% 
    pivot_longer(
      c(b_sjb_for, b_sjb_against),
      names_to = "cause",
      names_prefix = "b_sjb_",
      values_to = "b"
    ) %>%
    group_by(cause) %>% 
    median_qi(b) %>% 
    mutate(
      across(
        c(b, .lower, .upper), 
        f_num, digits = 2, retain.leading.zero = T
      ),
      b_text = glue(
        "$\\beta_{xy} = (b), [(.lower), (.upper)]$",
        .open = "(", .close = ")"
      )
    )

<!-- We also explored whether gender-related system justification predicted how acceptable participants judged protest actions for and against restricting abortion to be. To that end, we added two fixed effects to the model used in the preregistered analyses that estimated participants' responses as a function of their *z*-standardized system-justifying beliefs. As in Experiment 2, we found evidence for ideological asymmetry in this relationship: Participants who endorsed system-justifying beliefs considered the same protest action less acceptable when judging protests *against* restricting abortion (`r with(sjb, b_text[cause == "against"])`) but did not differ from participants who rejected system-justifying beliefs when judging protests *for* restricting abortion (`r with(sjb, b_text[cause == "for"])`). As in Experiment 2, our non-preregistered analyses thus suggested that system justification affects judgements about protests for a progressive cause (opposing abortion restrictions) but not for a conservative cause (supporting abortion restrictions). -->

@@

Our research was motivated by real-world examples that seemed to suggest that where people draw the line between acceptable and unacceptable protest actions depends on *who* the protesters are and *what* they are protesting---in other words, that there are double standards in judging collective action. Our findings, however, showed that it is not so simple. Contradicting our hypothesis of identity-based double standards, participants in two preregistered experiments did not show ingroup bias when judging controversial protest actions taken by ingroup and outgroup members to advance their group's interests. Instead, supporting our hypothesis of ideology-based double standards, participants judged the same controversial protest actions to be more acceptable when the protesters' cause aligned with their own (system-challenging) ideological position. 
<!-- This suggests that where observers draw the line between acceptable and unacceptable forms of collective action depends not so much on the protesters' and the observers' identities, but on the protesters' cause and how it aligns with the observers' political and ideological orientation.  -->
In the remainder of this General Discussion, we consider some limiting conditions on our findings and discuss theoretical and practical implications for understanding the often divided response to collective action.

Despite its strengths, our research is qualified by important limitations. First, our research examined the hypothesized relationships for three political causes---protecting workers' rights, defunding the police, restricting abortion---and thus provided only limited evidence that our findings generalize beyond those causes. Relying on few stimuli to establish an effect threatens the validity and replicability of research findings and constrains the generalizability of psychological research [@yarkoni_generalizability_2022]. Future research should sample a wider range of causes to address this pervasive but often ignored problem [@judd_treating_2012]. This is particularly important when studying collective action as recent theorizing [@jost_missing_2017] and research [@osborne_protesting_2019] highlighted the importance of differentiating between progressive and conservative causes. Second, our research was based on samples from two Western, Educated, Industrialized, Rich, and Democratic (WEIRD; Henrich et al., 2010) contexts---the United Kingdom and the United States---which limits the cross-cultural generalizability of our findings.

@@

Most broadly, our results showed that, for many controversial protest actions, there is no consensus on what is and is not an acceptable means to advance a cause. Instead, this judgement seems to depend on what the cause of a protest is and how it aligns with one's beliefs and values. In this way, our results confirm our contention that the distinction between normative and non-normative forms of collective action should be situated in the eye of the beholder. By developing an instrument for measuring where people draw the line between acceptable and unacceptable forms of collective action, we provided a paradigm that we hope will stimulate research on when and why people differ in making this distinction.

After all, this distinction can be consequential. If people believe that even extreme actions are acceptable means to advance a cause they support, especially if they moralize their support for that cause [@mooijman_moralization_2018], they might support violent extremism. In this way, our findings could explain why 45% of Republicans supported the storming of the U.S. Capitol on January 6, 2021 [@yougov_capitol_2021] even though it was illegal, destructive, and violent. If people believe that even non-violent actions are unacceptable means to advance a cause they oppose, they might support stifling dissent. In this way, our findings could explain why Republicans responded to Black Lives Matter protests by seeking to further criminalize disruptive protest [e.g., blocking highways, @pew_republicans_2021] that might be most effective at challenging injustice [@shuman_disrupting_2021; @shuman_protest_2022]. For those reasons, we  encourage researchers to use our paradigm to further investigate when and why people apply different standards when judging collective action for different causes.

@@

```{=latex}
\setcounter{table}{0}
\renewcommand{\thetable}{C\arabic{table}}

\section{Appendix C}

\caption{List of protest actions used in Experiment 3}
\small
\begin{tabularx}{\linewidth}{rXr}
\addlinespace
\toprule
\# & Action & Pr\\
\midrule
1 & make a public speech & 94\%\\
2 & hand out flyers, leaflets, or pamphlets & 92\%\\
3 & wear or display political symbols (e.g., patches, flags, bumper stickers) & 92\%\\
4 & hold meetings to influence the public & 92\%\\
5 & attend or organise a protest march & 91\%\\
6 & use social media (e.g., Facebook, Twitter, Instagram) to influence the public & 91\%\\
7 & donate to activist groups & 90\%\\
8 & refuse to buy goods or services from companies that advocate [against/for] restricting or banning abortion (boycott) & 90\%\\
9 & paste up posters with political messages in places where it is allowed and encouraged & 90\%\\
10 & join or form a group of activists & 90\%\\
11 & donate to politicians who advocate [for/against] restricting or banning abortion & 89\%\\
12 & pay for adverts on social media (e.g., Facebook, Twitter, Instagram) to influence public opinion & 84\%\\
13 & refuse to accept honors or awards in protest & 82\%\\
14 & refuse to work (strike) & 56\%\\
15 & visit people in their homes to convince them about the issue (canvassing, door knocking) & 50\%\\
16 & stand or sit in a building and refuse to leave (stand-in, sit-in) & 41\%\\
17 & attend a protest march even though it might be unlawful & 40\%\\
18 & \textit{protest outside the homes of politicians who advocate [against/for] restricting or banning abortion} & 40\%\\
19 & attend a protest march even though it might turn violent & 40\%\\
20 & paste up posters with political messages in places where it is not allowed or encouraged & 26\%\\
21 & enter and refuse to leave a building (occupation) & 25\%\\
22 & refuse to pay fees, fines, or taxes in protest & 23\%\\
23 & mock or insult individuals who are [against/for] restricting or banning abortion & 20\%\\
24 & spray paint political messages in public places & 15\%\\
25 & disrupt traffic (e.g., blocking roads) & 12\%\\
\bottomrule
\end{tabularx}
\caption*{\textit{Note.} Actions are ordered by the proportion of participants, across all conditions, who considered it to be an acceptable means of protest in Experiment 3. Action in \textit{italics} replaced an action used in Experiment 2 because it better fit the context of Experiment 3.}
